### 2018-10-05 figures/A1_iris.Rmd

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
# setwd("figures")
source("../scripts/01.5_loaddata.R")
```

```{r A1_iris, echo=FALSE, warning=FALSE, fig.width = 5.5, fig.height = 3}
ggplot() + 
  geom_point(data = iris, aes(x = Sepal.Length, y = Petal.Length))
```