### 2018-10-05 tables/01_iris_table.Rmd

---
output:
  pdf_document
header-includes: 
  - \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6)
```

```{r read_chunks, cache=FALSE, echo=FALSE}
knitr::read_chunk("../scripts/99_utils.R")
```

```{r source_utils, message=FALSE, results='hide', echo=FALSE, warning=FALSE}

```

```{r load_data, message=FALSE, results='hide', echo=FALSE, warning=FALSE}
source("../scripts/01.5_loaddata.R")
```

```{r iris_table, echo=FALSE, results='asis', message=FALSE, warning=FALSE}
knitr::kable(head(iris), caption = "Iris table")
```