### 2021-10-20 manuscript/figures.Rmd
read stan model from file

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

\pagebreak

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```