### 2021-10-21 manuscript/figures.Rmd

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: |
  \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

\pagebreak

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Censored lake area edf (solid line) and cdf estimate (dashed line).", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
```

```{r fig.cap="Stan model code", echo=FALSE, fig.width=1,out.height="5in"}
knitr::include_graphics("../manuscript/figures/stan.pdf")
```

```{r fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/bayesian_model-1.pdf")
```

```{r fig.cap="Median (black line) and central 95 percent interval estimates of total lake area (red lines). Here the true total lake area is marked with a dashed vertical line.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/bayesian_area-1.pdf")
```