### 2022-12-14 manuscript/figures.Rmd
more specific description of dashed line

---
title: "Figures"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: |
  \pagenumbering{gobble}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.path = "../figures/", fig.width = 5, fig.height = 6, dev = c("pdf"))
```

\pagebreak

```{r fig.cap="Realization of a Pareto (A) probability density function and (B) complementary cumulative distribution function compared against equivalent calculations on the HydroLAKES dataset.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/pareto_demo-1.pdf")
```

```{r fig.cap="Simulated lake area (blue line) where the portion (>= 0.1 $km^2$) calculated from the edf is denoted by a solid line and the portion (< 0.1 $km^2$) calculated from the transformed cdf is denoted by a dashed line. For comparison, the edf of the HydroLAKES dataset (red line) is also shown.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/predict_censor-1.pdf")
```

```{r fig.cap="Median (black line) and central 95 percent interval estimates of alpha (red lines). Here the 'true' alpha is 0.9 and is marked with a dashed line.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/bayesian_model-1.pdf")
```

```{r fig.cap="Median (black line) and central 95 percent interval estimates of total lake area (red lines). Here the true total lake area is marked with a dashed vertical line.", echo=FALSE, fig.width=1,out.height="4in"}
knitr::include_graphics("../manuscript/figures/bayesian_area-1.pdf")
```